CdGF_aucrf

### 2016-11-16 code/CdGF_AUCRF.Rmd
refine text of rmd

##Can we predict moribundity with the community structure at Day 0?

@@

om_loo_import_otus <- apply(raw_importance, 1, sum, na.rm = T)
om_loo_import_otus <- om_loo_import_otus[om_loo_import_otus > mean(om_loo_import_otus)]
important_otus_om_loo <- names(sort(om_loo_import_otus, decreasing = T))

#### plot rocs
# all mice
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='gray', lwd=2, add=T, lty=2)
plot(cv10f_roc, col='black', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(2,1,3), lwd=2, col=c('gray','black','blue'), bty='n')

#remove cage bias plots
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(om_roc, col='black', lwd=2, add=T, lty=1)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
plot(LOdO_roc, col='red', lwd=2, add=T, lty=3)
plot(om_loo_roc, col='blue', lwd=2, add=T, lty=2)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend(0.4, 0.125, legend=c(sprintf('One-mouse (AUC: %.3g)',om_roc$auc),
                           sprintf('Leave-1 source-out (AUC: %.3g)',LOcO_roc$auc),
                           sprintf('Leave-2 sources-out (AUC: %.3g)',LOdO_roc$auc),
                           sprintf('One-mouse-Leave-1 source-out (AUC: %.3g)',om_loo_roc$auc)),
       lty=c(1,3,3,2), lwd=2, col=c('black','purple','red','blue'), bty='n')
```

### Partial Dependency with Relative Abundance Overlayed Plots
###### Using optimal RF model generated from whole dataset
###### Plot labels contain MDA of each OTU

@@

Would adding initial increase/slope of CFU or day 1 CFU help predict?

* Adding CFU decreases AUC and improves Random model, likely due to samples lost due to lack of cfu data which are biased towards the severe samples
    - Day 1 CFU loses 2 samples  [AUCRF model with Day 1 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day1.html)
    - Day 2 cfu loses 14 samples, 10 are severe  [AUCRF model with Day 2 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day2.html)

Would it help to add community information such as diversity or enterotype?

* When adding enterotype, diversity, and CFU, invsimpson and cfu appear as features, but when cfu is removed, invsimpson is no longer a predictor

But how to show since model is already approaching 1?

* Rerun AUCRF with added features and see if they replace any OTUs

Leave 1 cage out RF with all otus gives and AUC of 0.70, but using the reduced features selected by AUCRF on the whole dataset gives an AUC of 0.80

To remove any cage effect RF could be picking out instead of the actual outcome, I first run RF on all OTUs to predict cage and separately human source. Then select the OTUS that are necessary for the lowest error rate (I went slightly above) from both cage and human souce RFs. Remove those OTUs from the OTU data frame and predict Euthanization with AUCRF

* Of note, I tried a few different seeds, and one otu (Turicibacter OTU 120) was identified as a classifier of cage/inocula but wasnt in others and was identified as a classifier of Euthanization. Also we occasionally lose OTU 50 as a classifier

####### Model run with both relative and absolute abundances and no significant difference in their output

>>>>

### Can we predict moribundity with the community structure at Day 0?

@@

om_loo_import_otus <- apply(raw_importance, 1, sum, na.rm = T)
om_loo_import_otus <- om_loo_import_otus[om_loo_import_otus > mean(om_loo_import_otus)]
important_otus_om_loo <- names(sort(om_loo_import_otus, decreasing = T))
```

#### RF with all mice input

```{r All data plt, fig.width=8, fig.height=8, results= 'hide'}
#### plot rocs
# all mice
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(otu_euth_roc, col='gray', lwd=2, add=T, lty=2)
plot(cv10f_roc, col='black', lwd=2, add=T, lty=1)
plot(LOO_roc, col='blue', lwd=2, add=T, lty=3)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('Out-Of-Bag (AUC: %.3g)',otu_euth_roc$auc),
                               sprintf('Leave-1 mouse-out (AUC: %.3g)',LOO_roc$auc),
                               sprintf('10-fold CV (AUC: %.3g)',cv10f_roc$auc) #,
#                               sprintf('OOB vs Leave-1-out: p=%.2g', roc.test(otu_euth_roc,LOO_roc)$p.value),
#                               sprintf('OOB vs 10-fold CV: p=%.2g', roc.test(otu_euth_roc,cv10f_roc)$p.value)
),lty=c(2,1,3), lwd=2, col=c('gray','black','blue'), bty='n')
```

#### RF with features from One_mouse_per_source

```{r one_mouse plots, fig.width=8, fig.height=8, results= 'hide'}
#remove cage bias plots
par(mar=c(4,4,1,1))
plot(c(1,0),c(0,1), type='l', lty=3, xlim=c(1.01,0), ylim=c(-0.01,1.01), xaxs='i', yaxs='i', ylab='', xlab='')
plot(om_roc, col='black', lwd=2, add=T, lty=1)
plot(LOcO_roc, col='purple', lwd=2, add=T, lty=3)
plot(LOdO_roc, col='red', lwd=2, add=T, lty=3)
plot(om_loo_roc, col='blue', lwd=2, add=T, lty=2)
mtext(side=2, text="Sensitivity", line=2.5)
mtext(side=1, text="Specificity", line=2.5)
legend('bottomright', legend=c(sprintf('One-mouse (AUC: %.3g)',om_roc$auc),
                           sprintf('Leave-1 source-out (AUC: %.3g)',LOcO_roc$auc),
                           sprintf('Leave-2 sources-out (AUC: %.3g)',LOdO_roc$auc),
                           sprintf('One-mouse-Leave-1 source-out (AUC: %.3g)',om_loo_roc$auc)),
       lty=c(1,3,3,2), lwd=2, col=c('black','purple','red','blue'), bty='n')
```

### Partial Dependency with Relative Abundance Overlayed Plots
###### Using features selected by one-mouse-per-source model

@@

#### Notes
* CFU data doesnt help model, likely due to samples lost due to lack of cfu data which are biased towards the severe samples  
    - Day 1 CFU loses 2 samples  [AUCRF model with Day 1 CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day1.html)  
    - Day 2 cfu loses 14 samples, 10 are severe  [AUCRF model with Day 2   CFU](~/Documents/Github/Schubert_humanCdGF_XXXX_2016/code/CdGF_AUCRF_withCFU_day2.html)  
* Diversity/Enterotype don't help model - When adding enterotype, diversity, and CFU, invsimpson and cfu appear as features, but when cfu is removed, invsimpson is no longer a predictor  
* Model run with both relative and absolute abundances and no significant difference in their output
